Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
clean up code and replace reported value with inline r code if possible

Effect \(E\) is derived from the Kullback-Leibler divergence (see
{[}appendix ref here{]}) formulated as

@@

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. These include the
distribution of the regressor and the type of the simulated model.

@@

Three experiments are conducted to investigate the difference between
conventional hypothesis testing and visual inference in the application
of linear regression diagnostics. The experiment I has ideal scenario
for conventional testing, where the visual test is not expected to
outperform the conventional test. Meanwhile, the experiment II is a
scenario where the conventional test is an approximate test, in which
the visual test may have a chance to match the performance of the
conventional test. The experiment III is designed for collecting human
responses to null lineups such that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
collect 7974 evaluations on 1152 uniqued lineups performed by 443
subjects throughout three experiments.

@@

As mentioned in Section \ref{calculating-p-values-for-the-visual-test},
\(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
estimated using {\textcolor{orange-red}{null}} lineups. Hence, 36
lineups with all combinations of \(n\) and \(X_{raw}\) and three
replications are included in experiment III. In these lineups, the data
of the data plot is generated from a model
{\textcolor{orange-red}{with zero effect size}}, while the data of the
19 null plots are generated using the same simulation method discussed
in Section \ref{sampling-from-the-null-distribution}.
{\textcolor{orange-red}{This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated from the null hypothesis. However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under the null hypothesis.}}
{\textcolor{orange-red}{To account for the fact that our simulation method for these lineups is not the Rorschach procedure,}}
we use the method suggested in \citet{vanderplas2021statistical} for
typical lineups containing a data plot to estimate \(\alpha\). We have
included a sensitivity analysis in the Appendix to examine the impact of
the variance of the \(\alpha\) estimate on our findings.

All lineups consist of only null plots are planned to be evaluated by 20
subjects. However, presenting only these lineups to subjects are
considered to be bad practices as subjects will lose interest quickly.
Therefore, we plan to collect 6 more evaluations on the 279 lineups with
\(X_{raw} \sim U(-1,1)\), result in
\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) / (20-2) = 133\)
subjects recruited for experiment III.

@@

There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and
133 subjects recruited for experiment I, II and III respectively. In the
total of 7974 lineup evaluations, 3744 use lineups produced by the
non-linearity model, 3510 use lineups produced by the heteroskedasticity
model, and 720 use null lineups. Besides, there are 886 attention checks
not included in the following analysis. The collated dataset is provided
in \texttt{vi\_survey} of the \texttt{visage} \texttt{R} package.

@@

For the non-linearity model, the power curve of RESET test climbs
aggressively from 13\% to around 70\% as \(log_e(E)\) increases from 0
to 2, while power of other tests respond inactively to the change of
effect, showing that RESET test is way more sensitive to the type of
model defects that being considered. Meanwhile, no noticeable visual
features can be spotted from the example residual plots.

@@

Surprisingly, the visual test rejects 1:2 out of c(NA, NA) (c(NA, NA)\%)
of lineups where the conventional test does not reject. Figure
\ref{fig:heter-example} shows this lineup. The data plot in position
seventeen displays a relatively strong heteroskedasticity pattern, and
has a strong effect size (\(log_e E=4.02\)). This is reflected by the
visual test \(p\text{-value} = 0.026\), but the Breusch-Pagan test
\(p\text{-value} = 0.056\), slightly above the significance cutoff of
\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
factors \(n=50\) (small sample size), \(\sigma=1\), and a uniform
distribution for the fitted values. It must be the small sample size
that may have resulted in the lack of detection.

>>>>

Effect \(E\) is derived from the Kullback-Leibler divergence (see
Appendix for the details) formulated as

@@

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. These include the
distribution of the regressor and the type of the simulated model
(modify this according to the result).

@@

Three experiments are conducted to investigate the difference between
conventional hypothesis testing and visual inference in the application
of linear regression diagnostics. The experiment I has ideal scenario
for conventional testing, where the visual test is not expected to
outperform the conventional test. Meanwhile, the experiment II is a
scenario where the conventional test is an approximate test, in which
the visual test may have a chance to match the performance of the
conventional test. The experiment III is designed for collecting human
responses to null lineups such that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
collect 7974 evaluations on 1152 unique lineups performed by 443
subjects throughout three experiments.

@@

As mentioned in Section \ref{calculating-p-values-for-the-visual-test},
\(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial} needs to be
estimated using null lineups. Hence, 36 lineups with all combinations of
\(n\) and \(X_{raw}\) and three replications are included in experiment
III. In these lineups, the data of the data plot is generated from a
model with zero effect size, while the data of the 19 null plots are
generated using the same simulation method discussed in Section
\ref{sampling-from-the-null-distribution}. This generation procedure
differs from the canonical Rorschach lineup procedure, which requires
that all 20 plots are generated from the null hypothesis. However, these
lineups serve the same fundamental purpose: to assess the number of
visually interesting plots generated under the null hypothesis.

To account for the fact that our simulation method for these lineups is
not the Rorschach procedure, we use the method suggested in
\citet{vanderplas2021statistical} for typical lineups containing a data
plot to estimate \(\alpha\). We have included a sensitivity analysis in
the Appendix to examine the impact of the variance of the \(\alpha\)
estimate on our findings.

All lineups consist of only null plots are planned to be evaluated by 20
subjects. However, presenting only these lineups to subjects are
considered to be bad practices as subjects will lose interest quickly.
Therefore, we plan to collect 6 more evaluations on the 279 lineups with
\(X_{raw} \sim U(-1,1)\), result in
\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
subjects recruited for experiment III.

@@

There are 2880, 2700 and \texttt{133\ *\ 18} lineups evaluation made by
160, 150 and 133 subjects recruited for experiment I, II and III
respectively. In the total of 7974 lineup evaluations, 3744 use lineups
produced by the non-linearity model, 4230 use lineups produced by the
heteroskedasticity model, and 720 use null lineups. Besides, there are
886 attention checks not included in the following analysis. The
collated dataset is provided in \texttt{vi\_survey} of the
\texttt{visage} \texttt{R} package.

@@

For the non-linearity model, the power curve of RESET test climbs
aggressively from 15\% to 69\% as \(log_e(E)\) increases from 0 to 2,
while power of other tests respond inactively to the change of effect,
showing that RESET test is way more sensitive to the type of model
defects that being considered. Meanwhile, no noticeable visual features
can be spotted from the example residual plots.

@@

Surprisingly, the visual test rejects 1 out of 33 (3\%) of lineups where
the conventional test does not reject. Figure \ref{fig:heter-example}
shows this lineup. The data plot in position seventeen displays a
relatively strong heteroskedasticity pattern, and has a strong effect
size (\(log_e E=4.02\)). This is reflected by the visual test
\(p\text{-value} = 0.026\), but the Breusch-Pagan test
\(p\text{-value} = 0.056\), slightly above the significance cutoff of
\(0.05\). This lineup was evaluated by 11 subjects, it has experimental
factors \(n=50\) (small sample size), \(\sigma=1\), and a uniform
distribution for the fitted values. It must be the small sample size
that may have resulted in the lack of detection.
